---
title: "SalesPrediction"
author: "Ankita Patidar"
date: "12/1/2019"
output:
  word_document: default
  pdf_document: default
---

```{r}
##install below packages if not already
##install.packages("cowplot")
##install.packages("caret")
##install.packages("dplyr")
#install.packages("data.table")
#install.packages("ggplot2")
```


```{r}
library(data.table)
#used for reading and manipulation
library(ggplot2)
#used for plotting
library(corrplot)
#used for correlation plotting
library(dplyr)
#used for manipulation and joining
library(cowplot)
# used for combining multiple plots
library(caret)
# used for modeling

```

## pre-processing

```{r}
#read the data

train = fread("BigMart_Train.csv")
test = fread("BigMart_Test.csv")
submission = fread("big_Mart_Submission.csv", skip = 1)

```


```{r}
#check the no. of columns and rows in train and test data
dim(train)
dim(test)
```


```{r}
#features of train and test data
names(train)
names(test)
```

Train dataset has 8523 rows and 12 features and test has 5681 rows and 11 columns. train has 1 extra column which is the target variable.
```{r}
#brief summary of the features of train and test data
str(train)
str(test)
```

```{r}
#combine train and test to perform modifcation on the data
test[,"Item_Outlet_Sales" := NA]
combi = rbind(train,test)
dim(combi)
str(combi)
```
## Item_MRP, Item_Visibility, Item_Outlet_Sales,Item_Weight is continous variables 
Visualize the continuous variables using histograms and categorical variables using bar plots.



```{r}
#Plot target variable "Item_Outlet_Sales"

ggplot(data = train) + geom_histogram(aes(train$Item_Outlet_Sales), binwidth = 100, fill = "darkgreen") + xlab("Item_Outlet_Sales")
# Target variable is right skewed

#Check the distribution of other numeric independent variable
p1 = ggplot(combi) + geom_histogram(aes(combi$Item_Weight), binwidth = 0.5, fill = "blue") + xlab("Item_Weight")
#No clear pattern in Item_Weight

p2 = ggplot(combi) + geom_histogram(aes(combi$Item_MRP), binwidth = 1, fill = "blue") + xlab("Item_MRP")
#4 different distributions for Item_MRP
p3 = ggplot(combi) + geom_histogram(aes(combi$Item_Visibility), binwidth = 0.005, fill = "blue") + xlab("Item_Visiblity")
#Item_Visibility is right-skewed

#plot_grid is used to combine plots
plot_grid(p1, p2,p3, nrow = 1)

```
##Outlet_Location_Type,Outlet_Type, Outlet_Identifier,Item_Type,Item_Identifier,Item_Fat_Content,Outlet_Size,

```{r}
#Plot for independent categorical variables

## Item_Fat_Content
ggplot(combi) + geom_bar(aes(combi$Item_Fat_Content), fill = "coral") + xlab("Item_Fat_Content")
#Combining LF, low fat, Low Fat into one Low Fat
combi$Item_Fat_Content[combi$Item_Fat_Content == "LF"] = "Low Fat"
combi$Item_Fat_Content[combi$Item_Fat_Content == "low fat"] = "Low Fat"
#combining reg and Regular into one Regular
combi$Item_Fat_Content[combi$Item_Fat_Content == "reg"] = "Regular"
ggplot(combi) + geom_bar(aes(combi$Item_Fat_Content), fill = "coral") + xlab("Item_Fat_Content")
```


```{r}
p4 = ggplot(combi %>% group_by(Item_Type) %>% summarise(Count =n())) + geom_bar(aes(Item_Type, Count), stat = "identity", fill = "coral") + xlab("") + geom_label(aes(Item_Type,  Count, label =  Count), vjust = 0.5) + theme(axis.text.x = element_text(angle= 45, hjust = 1)) + ggtitle("Item_Type")


p5 = ggplot(combi %>% group_by(Item_Identifier) %>% summarise(Count =n())) + geom_bar(aes(Item_Identifier, Count), stat = "identity", fill = "coral") + xlab("") + geom_label(aes(Item_Identifier,  Count, label =  Count), vjust = 0.5) + theme(axis.text.x = element_text(angle= 45, hjust = 1)) + ggtitle("Item_Identifier")


p6 = ggplot(combi %>% group_by(Outlet_Size) %>% summarise(Count =n())) + geom_bar(aes(Outlet_Size, Count), stat = "identity", fill = "coral") + xlab("") + geom_label(aes(Outlet_Size,  Count, label =  Count), vjust = 0.5) + theme(axis.text.x = element_text(angle= 45, hjust = 1)) + ggtitle("Outlet_Size")
p6


p7 = ggplot(combi %>% group_by(Outlet_Identifier) %>% summarise(Count =n())) + geom_bar(aes(Outlet_Identifier, Count), stat = "identity", fill = "coral") + xlab("") + geom_label(aes(Outlet_Identifier,  Count, label =  Count), vjust = 0.5) + theme(axis.text.x = element_text(angle= 45, hjust = 1)) + ggtitle("Outlet_Identifier")

second_row = plot_grid( p6, p7, nrow =  1)

plot_grid(p4, second_row, ncol = 1)

```


```{r}
p8 = ggplot(combi %>% group_by(Outlet_Establishment_Year) %>% summarise(Count = n())) +   geom_bar(aes(factor(Outlet_Establishment_Year), Count), stat = "identity", fill = "coral1") +  geom_label(aes(factor(Outlet_Establishment_Year), Count, label = Count), vjust = 0.5) +  xlab("Outlet_Establishment_Year") +  theme(axis.text.x = element_text(size = 8.5))
#1998 has lesser number of observation


p9 = ggplot(combi %>% group_by(Outlet_Type) %>% summarise(Count = n())) +   geom_bar(aes(Outlet_Type, Count), stat = "identity", fill = "coral1") +  geom_label(aes(factor(Outlet_Type), Count, label = Count), vjust = 0.5) +  theme(axis.text.x = element_text(size = 8.5))
#Supermarket Type 1 seems to be most popular

plot_grid(p8, p9, ncol = 2)
```


```{r}
#bi-variate analysis - explore the independent variables with respect to the target variable.
#use of scatter plots for the continuous or numeric variables and violin plots for the categorical variables.
# extracting train data from the combined data'

train = combi[1:nrow(train)]

# Item_Weight vs Item_Outlet_Sales

p10 = ggplot(train) + geom_point(aes(Item_Weight, Item_Outlet_Sales), colour = "red", alpha = 0.3) +theme(axis.title = element_text(size=8.5))
p10

# Item_Visibility vs Item_Outlet_Sales

p11 = ggplot(train) + geom_point(aes(Item_Visibility,Item_Outlet_Sales), colour = "red", alpha = 0.3) + theme(axis.title = element_text(size = 8.5))
p11 
#there is a string of points at Item_Visibility = 0.0 which seems strange as item visibility cannot be completely zero

# Item_MRP vs Item_Outlet_Sales

p12 = ggplot(train) + geom_point(aes(Item_MRP, Item_Outlet_Sales), colour = "violet", alpha = 0.3) +  theme(axis.title = element_text(size = 8.5))
p12
```


```{r}
# Item_Type vs Item_Outlet_Sales 

p13 = ggplot(train) + geom_violin(aes(Item_Type,Item_Outlet_Sales), fill = "Magenta") + theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.text = element_text(size = 6), axis.title = element_text(size = 8.5))
p13

# Item_Fat_Content vs Item_Outlet_Sales
p14 = ggplot(train) +       geom_violin(aes(Item_Fat_Content, Item_Outlet_Sales), fill = "magenta") +      theme(axis.text.x = element_text(angle = 45, hjust = 1),            axis.text = element_text(size = 8),            axis.title = element_text(size = 8.5))

  p14

# Outlet_Identifier vs Item_Outlet_Sales

p15= ggplot(train) +       geom_violin(aes(Outlet_Identifier, Item_Outlet_Sales), fill = "magenta") +      theme(axis.text.x = element_text(angle = 45, hjust = 1),            axis.text = element_text(size = 8),            axis.title = element_text(size = 8.5))

p15

p16 = ggplot(train) + geom_violin(aes(Outlet_Size, Item_Outlet_Sales), fill = "magenta")

p16
#The distribution of ‘Small’ Outlet_Size is almost identical to the distribution of the blank category (first vioin) of Outlet_Size. So, we can substitute the blanks in Outlet_Size with ‘Small’.
```

```{r}
p17 = ggplot(train) + geom_violin(aes(Outlet_Location_Type, Item_Outlet_Sales), fill = "blue")
#Tier 1 and Tier 3 locations of Outlet_Location_Type look similar.
p18 = ggplot(train) + geom_violin(aes(Outlet_Type, Item_Outlet_Sales), fill = "blue")
#Grocery Store has most of its data points around the lower sales

p17
p18
```


```{r}
#finding missing values in variables
sum(is.na(combi$Item_Weight))
#2439 missing values in Item_Weight
sum(is.na(combi$Item_Fat_Content))
sum(is.na(combi$Item_Identifier))
sum(is.na(combi$Item_Visibility))
sum(is.na(combi$Item_Type))
sum(is.na(combi$Item_MRP))
sum(is.na(combi$Outlet_Identifier))
sum(is.na(combi$Outlet_Establishment_Year))
sum(is.na(combi$Outlet_Size))
sum(is.na(combi$Outlet_Location_Type))
sum(is.na(combi$Outlet_Type))
sum(is.na(combi$Item_Outlet_Sales))
#5681 missing values in Item_Outlet_Sales
```
```{r}
#impute Item_Weight with mean weight based on the Item_Identifier variable
missing_index = which(is.na(combi$Item_Weight))
for (i in missing_index){
  item = combi$Item_Identifier[i]
  combi$Item_Weight[i] = mean(combi$Item_Weight[combi$Item_Identifier == item], na.rm = T)
}

sum(is.na(combi$Item_Weight))

```



```{r}
ggplot(combi) + geom_histogram(aes(Item_Visibility), bins =100)
zero_index = which(combi$Item_Visibility == 0)
#zeroes in Item_Visibility variable can be replaced with Item_Identifier wise mean values of Item_Visibility
for (i in zero_index){
  item = combi$Item_Identifier[i]
  combi$Item_Visibility[i] = mean(combi$Item_Visibility[combi$Item_Identifier==item], na.rm = T)
}

ggplot(combi) + geom_histogram(aes(Item_Visibility), bins = 100)
#the issue of zero item visibility has been resolved.
```

Create new features to help improving the model performance:

Item_Type_new: Broader categories for the variable Item_Type.
Item_category: Categorical variable derived from Item_Identifier.
Outlet_Years: Years of operation for outlets.
price_per_unit_wt: Item_MRP/Item_Weight
Item_MRP_clusters: Binned feature for Item_MRP.


```{r}
#the Item_Type variable can be classified into the categories perishable and non_perishable
perishable = c("Breads", "Breakfast", "Dairy", "Fruits and Vegetables", "Meat", "Seafood")

non_perishable = c("Baking Goods", "Canned", "Frozen Foods", "Hard Drinks", "Health and Hygiene", "Household", "Soft Drinks")

# create a new feature 'Item_Type_new' 

combi[, Item_Type_New := ifelse(Item_Type %in% perishable, "perishable", ifelse(Item_Type %in% non_perishable, "non_perishable", "not sure"))]
```


```{r}
#the first 2 characters of Item_Identifier, i.e., ‘DR’, ‘FD’, and ‘NC’ stand for drinks, food, and non-consumable.
table(combi$Item_Type, substr(combi$Item_Identifier,1,2))
```

```{r}
combi[, Item_Category := substr(combi$Item_Identifier,1,2)]

combi$Item_Fat_Content[combi$Item_Category == "NC"] =  "non-edible"


# years of operation for outlets 
combi[, Item_Years := 2013 - combi$Outlet_Establishment_Year]
combi$Outlet_Establishment_Year = as.factor(combi$Outlet_Establishment_Year)

# Price per unit weight 

combi[, PPUW := Item_MRP/Item_Weight]

# creating new independent variable - Item_MRP_clusters
combi[,Item_MRP_Clusters := ifelse(Item_MRP  < 69, "1st", ifelse(Item_MRP >= 69 & Item_MRP < 139, "2nd", ifelse(Item_MRP >= 139 & Item_MRP < 203, "3rd", "4th")))]
```


```{r}
#Label encode Outlet_Size and Outlet_Location_Type as these are ordinal variables.
combi[,Outlet_Size_Num := ifelse(Outlet_Size == "Small", 0, ifelse(Outlet_Size == "Medium", 1,2))]

combi[,Outlet_Location_Type_num := ifelse(Outlet_Location_Type == "Tier 3", 0,                                          ifelse(Outlet_Location_Type == "Tier 2", 1, 2))]

```

```{r}
# removing categorical variables after label encoding 

combi[,c("Outlet_Size", "Outlet_Location_Type") := NULL]
```



```{r}
#One hot encoding for the categorical variable
ohe = dummyVars("~." , data = combi[, -c("Item_Identifier", "Outlet_Establishment_Year", "Item_Type")], fullRank = T)

ohe_df = data.table(predict(ohe, combi[,-c("Item_Identifier", "Outlet_Establishment_Year", "Item_Type")]))

combi = cbind(combi[,"Item_Identifier"], ohe_df)
```



```{r}
combi[,Item_Visibility := log(Item_Visibility + 1)]
# log + 1 to avoid division by zero
combi[,PPUW := log(PPUW + 1)]
```

```{r}
#scale and center the numeric variables to make them have a mean of zero, standard deviation of one and scale of 0 to 1.

# index of numeric features
num_vars = which(sapply(combi, is.numeric))

num_vars_names = names(num_vars)
combi_numeric = combi[,setdiff(num_vars_names, "Item_Outlet_Sales"), with= F]
prep_num = preProcess(combi_numeric, method = c("center", "scale"))
combi_numeric_norm = predict(prep_num, combi_numeric)

 # removing numeric independent 
combi[,setdiff(num_vars_names, "Item_Outlet_Sales") := NULL] 

combi = cbind(combi, combi_numeric_norm)

#Splittng

train = combi[1:nrow(train)] 
test = combi[(nrow(train) + 1):nrow(combi)]
test[,Item_Outlet_Sales := NULL] 
# removing Item_Outlet_Sales as it contains only NA for test dataset
```


```{r}
cor_train = cor(train[,-c("Item_Identifier")])
corrplot(cor_train, method = "circle", type = "lower", tl.cex = 0.9)
```


```{r}
#Regression Model
linear_reg_model = lm(Item_Outlet_Sales ~., data =train[,-c("Item_Identifier")])
linear_reg_model

```

```{r}
# preparing dataframe for submission and writing it in a csv file 
submission$Item_Outlet_Sales = predict(linear_reg_model, test[,-c("Item_Identifier")]) 
write.csv(submission, "Linear_Reg_submit.csv", row.names = F)
```







